\section{Conclusions}
\label{sec:concl}
 
We considered DAG-SyDSs, that is,
synchronous dynamical systems over directed acyclic graphs.
We established several properties of the phase spaces of
such dynamical systems and used those properties in establishing
complexity results and developing algorithms 
for problems such as reachability and convergence guarantee.
There are several directions for future research.
For example, it will be of interest to consider 
DAG-SyDSs where local functions are from other classes
(e.g., linear functions, weighted threshold functions)
and establish properties of phase spaces.
Similarly, one may also consider restrictions
on the graph structure (e.g., dags where nodes have
bounded in-degrees) and investigate phase space properties
as well as the complexity of reachability and similar problems.
As another extension, it is also of interest to investigate
synchronous dynamical systems over directed graphs which are
near-acylic in some fashion (e.g., removing a few directed
edges leads to a dag).


\iffalse
%%%%%%%%%%%%%%%%%%%%%%%%%%
\noindent
\paragraph{Conclusions.}


\paragraph{Open Questions.}

\begin{itemize}
\item For DAG linear SyDSs, how long can a phase space cycle and transient be,
as a function of depth.

\item For bounded degree DAG  SyDSs, how long can a phase space cycle and transient be,
as a function of depth.

\item For various classes of Boolean functions of interest, 
how long can a phase space cycle and transient be?
(If they are polynomially bounded as a function of the number of nodes,
then reachability is polynomially solvable.)
\end{itemize}
%%%%%%%%%%%%%%%%%%%%%%%%%%
\fi
