\section{Reachability Problem for Special Classes of DAG-SyDSs}
\label{sec:reach_special}

\subsection{DAG-SyDSs with Monotone Local Functions}
\label{sse:monotoner_functions}

The class of monotone Boolean functions were defined in
Section~\ref{sec:defs}.
A monotone DAG-SyDS is a DAG-SyDS in which every local
function is monotone.
In this section, we show that the reachability problem
for monotone DAG-SyDSs can be solved efficiently.
This result is based on the following theorem. 

\begin{theorem}\label{thm:monotone_stable}
For a DAG monotone SyDS, every cycle is a fixed point, 
and the length of any transient does not exceed the number of levels.
\end{theorem}

\noindent
\textbf{Proof:}~ See supplement.

\iffalse
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\noindent
\textbf{Proof:}~ 
By induction on the number of levels for each level $i$, 
each level $i$ node is stable after at most $i$ steps.

From Lemma \ref{lem:level_zero_nodes}, every level 0 node is either
stable or alternating  at time 1.  The complement function is not
monotone, so a  level 0 node is cannot be alternating, and thus is
stable at time 1.

Suppose that all the incoming edges to a given node $v$ have stable
values at time $t$.  From Lemma \ref{lem:all_inputs_stable}, node
$v$ is either stable or alternating  at time $t+1$.  Since the local
transition function for node $v$ is monotone, and all the incoming
edges have stable values at time $t$, node $v$ is stable at time
$t+1$. \QED

We observe that that for every $L \geq 0$, there exists a DAG
monotone SyDS with a transient of length $L$.  Consider the SyDS
whose underlying graph is a directed chain of $L$ nodes.  The local
transition function of the level 0 node is the constant 1, and of
every other node is the {\em or} of its value and the value from
the incoming edge.  The configuration of all zeros takes $L$ steps
to reach the fixed point of all ones.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\fi

Since Theorem~\ref{thm:monotone_stable} provides a polynomial bounds
on the lengths of transients and cycles in the phase space of any
monotone DAG-SyDS, the following corollary is immediate.

\begin{corollary}\label{cor:poly_monotone_dag_syds}
The Reachability problem is efficiently solvable for monotone
DAG-SyDSs.
\end{corollary}


